---
title: "Systolic Blood Pressure"
author: "Can Li"
date: "`r format(Sys.Date())`"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.align = "center",
                      dev = "png",
                      cache = TRUE)
```

### **Introduction**

> This report analyzes blood pressure data from the survey collected by the United States National Center for Health Statistics (NCHS). This center has conducted a series of health and nutrition surveys since the 1960’s.

> Starting in 1999, about 5,000 individuals of all ages have been interviewed every year and then they complete the health examination component of the survey. Part of this dataset is made available via the **NHANES** package.

### **Libraries**
```{r loading-libs}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("NHANES")
library("tidyverse")
library("NHANES")
```

### **Datasets**
```{r loading-data}
data(NHANES)
```
### **Summary of bp data by Age Group**
```{R group_gender, echo=2}
bp_gender <- NHANES %>% filter(!(is.na(AgeDecade)|is.na(BPSysAve))) %>% 
  group_by(Gender) %>% 
  arrange(Gender, AgeDecade) %>% 
  select(Gender, AgeDecade, BPSysAve)

bp_gender %>% group_by(AgeDecade) %>% 
    summarize(average=mean(BPSysAve), sd=sd(BPSysAve), 
              min = min(BPSysAve), max = max(BPSysAve))
```

### **Summary of bp data by Age Group and Gender**
```{r summary, echo=FALSE}
## Summary of mean, sd, min, max
bp_summary <- NHANES %>% filter(!(is.na(AgeDecade)|is.na(BPSysAve))) %>% 
  group_by(AgeDecade, Gender) %>% 
  summarize(average=mean(BPSysAve), sd=sd(BPSysAve), 
            min = min(BPSysAve), max = max(BPSysAve))
bp_summary
```

### **A Boxplot to Compare bp between Gender in Age Groups**
```{r gender_diff}
bp_gender %>% ggplot(aes(AgeDecade, BPSysAve, color= Gender)) +
  geom_boxplot()
```

### **Summary of bp by race as reported in the Race1 variable**
```{r race-diff}
bp_race <- NHANES %>% 
  filter(Gender == "male" & AgeDecade == " 40-49" & !is.na(BPSysAve)) %>% 
  group_by(Race1) %>% select(Race1, BPSysAve) 

bp_race %>% ggplot(aes(Race1, BPSysAve, color=Race1)) +
  geom_boxplot()

tab_race <- bp_race %>% 
  summarize(average = mean(BPSysAve), sd = sd(BPSysAve)) %>% arrange(average)
tab_race   
```




